---
title: "Final Project"
author: "Jacie Wright"
date: "4/17/2022"
output: html_document
---

```{r setup, include=FALSE}
#install.packages(tidyverse)
library(knitr)
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```

```{r}

## This code block reads and re-writes all of the data into "dataFrame" to be used in each analysis

data <- read.csv('statsResponses v0.csv')
dataFrame <- data
  #GPA "success" range: 5 = Excellent, 4 = Good, 3 = Moderate, 2 = Poor, 1 = Bad
  dataFrame$GPA[dataFrame$GPA == "3.5+"] = 5
  dataFrame$GPA[dataFrame$GPA == "3-3.4"] = 4
  dataFrame$GPA[dataFrame$GPA == "2.5-2.9"] = 3
  dataFrame$GPA[dataFrame$GPA == "2-2.4"] = 2
  dataFrame$GPA[dataFrame$GPA == "<1.9"] = 1
  dataFrame$GPA = as.integer(dataFrame$GPA)
  #Modify transfer GPAs
  dataFrame$Transfers[dataFrame$Transfers == "3.5+"] = 5
  dataFrame$Transfers[dataFrame$Transfers == "3-3.4"] = 4
  dataFrame$Transfers[dataFrame$Transfers == "2.5-2.9"] = 3
  dataFrame$Transfers[dataFrame$Transfers == "2-2.4"] = 2
  dataFrame$Transfers[dataFrame$Transfers == "<1.9"] = 1
 
  #Modify credits
  dataFrame$Credits[dataFrame$Credits == "16+"] = 16
  dataFrame$Credits = as.integer(dataFrame$Credits)
  
  #Modify concentration; no struggle concentrating (1) to struggle concentrating every day (4)
  dataFrame$Concentration[dataFrame$Concentration == "Not at all"] = "1"
  dataFrame$Concentration[dataFrame$Concentration == "Several days"] = "2"
  dataFrame$Concentration[dataFrame$Concentration == "More than half of the days"] = "3"
  dataFrame$Concentration[dataFrame$Concentration == "Nearly every day"] = "4"
  dataFrame$Concentration = as.integer(dataFrame$Concentration)
   
  #Work Overall Part 1; I love what I'm doing (1) to I hate what I'm doing (5)
  #dataFrame$Work.Overall.Part.1[dataFrame$Work.Overall.Part.1 ==]
  
  #Work OVerall Part 2; my job will help my future (1) my job will not help my future (5)
  
  
  #Hours of sleep; 
  dataFrame$Hours.of.Sleep[dataFrame$Hours.of.Sleep == "0-2"] = "1"
  dataFrame$Hours.of.Sleep[dataFrame$Hours.of.Sleep == "3-4"] = "2"
  dataFrame$Hours.of.Sleep[dataFrame$Hours.of.Sleep == "5-6"] = "3"
  dataFrame$Hours.of.Sleep[dataFrame$Hours.of.Sleep == "7-8"] = "4"
  dataFrame$Hours.of.Sleep[dataFrame$Hours.of.Sleep == "8+"] = "5"
  
  #Struggle to sleep
  
  #Activity level; no exercise (1) to nearly every day in a week (4)
  dataFrame$Activities[dataFrame$Activities == "0 days a week"] = "1"
  dataFrame$Activities[dataFrame$Activities == "1-2 days a week"] = "2"
  dataFrame$Activities[dataFrame$Activities == "3-4 days a week"] = "3"
  dataFrame$Activities[dataFrame$Activities == "5-7 days a week"] = "4"
  dataFrame$Activities = as.integer(dataFrame$Activities)
  
  #Activity durations; need a conditional check where if activity level = 1 set to 0
  
  

```

## Introduction (Jc)

## Data Collection

### Limitations of collection

## Data and Analysis

### Limitations of Analysis


``` {r}

plot(dataFrame$GPA, type = "l", col = "red", xlim = c(0,length(dataFrame$Timestamp)), ylim = c(0,max(dataFrame$Hours.of.Studying)))
lines(dataFrame$Hours.of.Studying, col = "black")

```


``` {r}

######## Tony added code 04/22/2022 In-Progress ###############

#Modify hours of sleep to be numeric
dataFrame$Hours.of.Sleep = as.integer(dataFrame$Hours.of.Sleep)

#Create list of data without NAs in commute times
data_no_na_commute_times = dataFrame %>% drop_na(Commute.Time)


#lines(x= 0:100, y = dunif(seq(0, 1, 0.01), min=0, max=50), col="green")
#lines(1.35*dnorm(0:100, mean(temp_data_row_means), sd(temp_data_row_means)), col='purple')
#lines(3.5*dpois(0:100, mean(temp_data_row_means)), col='red')
#lines(7*dexp(0:100, 1/mean(temp_data_row_means)), col='blue')
#lines(3.5*dgamma(0:100, mean(temp_data_row_means)), col='orange')

#legend(48, 0.25, legend=c("Uniform Dist", "Normal Dist", "Poisson Dist", "Exponential Dist", "Gamma Dist"),
       #col=c("green", "purple", "red", "blue", "orange"), lty=1, cex=0.75)

########################### VIEW THE RELATIONSHIP BETWEEN EXERCISE AND GPA ##############


#Create data
set.seed(1)
 
# Base histogram
hist(dataFrame$Activities,  col=rgb(1,0,0,0.5), xlab="Levels of Exercise and GPA", 
     ylab="Response count", main="" )

# Second with add=T to plot on top
hist(dataFrame$GPA, col=rgb(0,0,1,0.5), add=T)

# Add legend
legend("topright", legend=c("Exercise","GPA"), col=c(rgb(1,0,0,0.5), 
     rgb(0,0,1,0.5)), pt.cex=2, pch=15 )


# Create scatter plot for with GPA as y-axis and Exercise as X-value

#I count the occurence of each couple of values. Eg : number of time a=1 and b=1, number of time a=1 and b=2 etc...
AA <- xyTable(dataFrame$GPA,dataFrame$Activities)
 
#Now I can plot this ! I represent the dots as big as the couple occurs often
coeff_bigger <- 2
plot(AA$x , AA$y , cex=AA$number*coeff_bigger  , pch=16 , col=rgb(0,0,1,0.5) , xlab= "Gpa levels" , ylab="Exercise levels" , xlim=c(0,6) , ylim=c(0,6), main="Excercse levels per week VS GPA")
text(AA$x , AA$y , AA$number )
 
#Note : It's easy to make a function that will compute this kind of plot automaticaly :
represent_discrete_variable <- function(var1, var2 , coeff_bigger){
  AA=xyTable(var1,var2)
  plot(AA$x , AA$y , cex=AA$number*coeff_bigger  , pch=16 , col="chocolate1" , xlab= "value of a" , ylab="value of b" )
  text (AA$x , AA$y , AA$number )
}


########## Comparison for subquestion of exercise levels and energy levels ##############################

# Base histogram
hist(dataFrame$Energy.Level,  col=rgb(0,1,0,0.5), xlab="Levels of Exercise and Energy", 
     ylab="Response Count", main="Exercise per Week vs Energy Levels of Last Two Weeks", ylim = c(0,25))

# Second with add=T to plot on top
hist(dataFrame$Activities, col=rgb(1,0,0,0.5), add=T)

# Add legend
legend("topright", legend=c("Energy Levels","Exercise"), col=c(rgb(0,1,0,0.5), 
     rgb(1,0,0,0.5)), pt.cex=2, pch=15 )

```
```{r}
########## Hypothesis Test for Exercise & GPA vs No Exercise and GPA ##############################

no_exercise_GPAs = dataFrame$GPA[dataFrame$Activities == 1]
exercise_GPAs = dataFrame$GPA[dataFrame$Activities != 1]

#For the t.test we want to make the claim "Exercise of any frequency raises a student GPA". This mean that our null hypothesis we must claim that the opposite is true or that "There is no difference in GPAs" is true.

#If we run a t.test with 0.95 confidence level, our p-value must be lower than 0.05 for us to refuse the null hypothesis that "There is no difference in GPAs" is true.
#Also, if the confidence interval holds the value of 0.0 then this means that our difference in means from both sets are negligible. 

t.test(no_exercise_GPAs, exercise_GPAs, conf.level = 0.95)


```


### Exercise Frequency vs GPA

When first looking at the data on exercise levels vs academic success, our first thoughts were leaning toward seeing GPA growth with a higher frequency of exercise. It’s normal to think that some exercise would help on numerous levels than having no exercise at all but in order to get a better analysis of this topic, we decided to create a few items that will help in our analysis. First, let us look at the scatter plot above. This scatter plot represents the frequency of student responses to their weekly frequency of exercise days with level one being zero days, level two being one to two days, level three being three to four days, and level four being five to seven days. A similar level and range scheme can be seen for the values of GPA data with level one being less than one-point-nine, level two being two to two-point-four, level three being two-point-five to two-point-nine, level four being three-point-zero to three-point-four, and lastly, level five being three-point-five and above.

When looking at the scatter plot we can see that there may be a correlation between a higher GPA and a higher frequency of exercising per week. The main thing that stands out here is that GPA level five (being a three-point-five GPA and above) seems to hold a little over half of the number of survey responses gathered. This may be an indication that exercise helps increase GPA but we must do further analysis to have strong supporting evidence. The next step we can take is to do a hypothesis test with two sample means. Specifically, the two-sample means of the group of students who do not exercise during the week and the sample means of the group of students who do exercise during the week. The claim for this hypothesis test will be that the two sample means of students who do and do not exercise differ. This would then mean that our null hypothesis must be that the means of students who do and do not exercise are equal. Also, we will run the hypothesis test with a 95% confidence level.

The results of the hypothesis test show a p-value of 0.9493. In order for us to accept the null hypothesis, we must get a resulting p-value that is less than 0.05. Seeing that ours is far greater than this means that we must fail to reject the null hypothesis. Adding to this, we can see that the confidence interval result of -0.3915901 to 0.3675300 includes the value of 0.0 which shows that there was no significant difference between the means of students who do and do not exercise. Although we have the combination of needing to refuse the null hypothesis and our confidence interval containing the value of 0.0, all we can truly say with these results is that we must refuse the null hypothesis and this essentially just lets us know the null hypothesis of our claim is not supported by our current data. This in turn points to the fact that further research is needed to answer the question “Does exercise increase academic success?” as our current sample data shows no significant difference in the means of GPAs of students who do and students who do not exercise.