---
title: "Final Project"
author: "Jacie Wright"
date: "4/17/2022"
output: html_document
---

```{r setup, include=FALSE}
#install.packages(tidyverse)
library(knitr)
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```

```{r}

## This code block reads and re-writes all of the data into "dataFrame" to be used in each analysis

data <- read.csv('statsResponses v0.csv')
dataFrame <- data
  #GPA "success" range: 5 = Excellent, 4 = Good, 3 = Moderate, 2 = Poor, 1 = Bad
  dataFrame$GPA[dataFrame$GPA == "3.5+"] = 5
  dataFrame$GPA[dataFrame$GPA == "3-3.4"] = 4
  dataFrame$GPA[dataFrame$GPA == "2.5-2.9"] = 3
  dataFrame$GPA[dataFrame$GPA == "2-2.4"] = 2
  dataFrame$GPA[dataFrame$GPA == "<1.9"] = 1
  dataFrame$GPA = as.integer(dataFrame$GPA)
  #Modify transfer GPAs
  dataFrame$Transfers[dataFrame$Transfers == "3.5+"] = 5
  dataFrame$Transfers[dataFrame$Transfers == "3-3.4"] = 4
  dataFrame$Transfers[dataFrame$Transfers == "2.5-2.9"] = 3
  dataFrame$Transfers[dataFrame$Transfers == "2-2.4"] = 2
  dataFrame$Transfers[dataFrame$Transfers == "<1.9"] = 1
 
  #Modify credits
  dataFrame$Credits[dataFrame$Credits == "16+"] = 16
  dataFrame$Credits = as.integer(dataFrame$Credits)
  
  #Modify concentration; no struggle concentrating (1) to struggle concentrating every day (4)
  dataFrame$Concentration[dataFrame$Concentration == "Not at all"] = "1"
  dataFrame$Concentration[dataFrame$Concentration == "Several days"] = "2"
  dataFrame$Concentration[dataFrame$Concentration == "More than half of the days"] = "3"
  dataFrame$Concentration[dataFrame$Concentration == "Nearly every day"] = "4"
  dataFrame$Concentration = as.integer(dataFrame$Concentration)
   
  #Work Overall Part 1; I love what I'm doing (1) to I hate what I'm doing (5)
  #dataFrame$Work.Overall.Part.1[dataFrame$Work.Overall.Part.1 ==]
  
  #Work OVerall Part 2; my job will help my future (1) my job will not help my future (5)
  
  
  #Hours of sleep; 
  dataFrame$Hours.of.Sleep[dataFrame$Hours.of.Sleep == "0-2"] = "1"
  dataFrame$Hours.of.Sleep[dataFrame$Hours.of.Sleep == "3-4"] = "2"
  dataFrame$Hours.of.Sleep[dataFrame$Hours.of.Sleep == "5-6"] = "3"
  dataFrame$Hours.of.Sleep[dataFrame$Hours.of.Sleep == "7-8"] = "4"
  dataFrame$Hours.of.Sleep[dataFrame$Hours.of.Sleep == "8+"] = "5"
  
  #Struggle to sleep
  
  #Activity level; no exercise (1) to nearly every day in a week (4)
  dataFrame$Activities[dataFrame$Activities == "0 days a week"] = "1"
  dataFrame$Activities[dataFrame$Activities == "1-2 days a week"] = "2"
  dataFrame$Activities[dataFrame$Activities == "3-4 days a week"] = "3"
  dataFrame$Activities[dataFrame$Activities == "5-7 days a week"] = "4"
  dataFrame$Activities = as.integer(dataFrame$Activities)
  
  #Activity durations; need a conditional check where if activity level = 1 set to 0
  
  

```

## Introduction (Jc)

## Data Collection

### Limitations of collection

## Data and Analysis

### Limitations of Analysis


``` {r}

plot(dataFrame$GPA, type = "l", col = "red", xlim = c(0,length(dataFrame$Timestamp)), ylim = c(0,max(dataFrame$Hours.of.Studying)))
lines(dataFrame$Hours.of.Studying, col = "black")

```


``` {r}

######## Tony added code 04/22/2022 In-Progress ###############

#Modify hours of sleep to be numeric
dataFrame$Hours.of.Sleep = as.integer(dataFrame$Hours.of.Sleep)

#Create list of data without NAs in commute times
data_no_na_commute_times = dataFrame %>% drop_na(Commute.Time)


#lines(x= 0:100, y = dunif(seq(0, 1, 0.01), min=0, max=50), col="green")
#lines(1.35*dnorm(0:100, mean(temp_data_row_means), sd(temp_data_row_means)), col='purple')
#lines(3.5*dpois(0:100, mean(temp_data_row_means)), col='red')
#lines(7*dexp(0:100, 1/mean(temp_data_row_means)), col='blue')
#lines(3.5*dgamma(0:100, mean(temp_data_row_means)), col='orange')

#legend(48, 0.25, legend=c("Uniform Dist", "Normal Dist", "Poisson Dist", "Exponential Dist", "Gamma Dist"),
       #col=c("green", "purple", "red", "blue", "orange"), lty=1, cex=0.75)

########################### VIEW THE RELATIONSHIP BETWEEN EXERCISE AND GPA ##############


#Create data
set.seed(1)
 
# Base histogram
hist(dataFrame$Activities,  col=rgb(1,0,0,0.5), xlab="Levels of Exercise and GPA", 
     ylab="Response count", main="" )

# Second with add=T to plot on top
hist(dataFrame$GPA, col=rgb(0,0,1,0.5), add=T)

# Add legend
legend("topright", legend=c("Exercise","GPA"), col=c(rgb(1,0,0,0.5), 
     rgb(0,0,1,0.5)), pt.cex=2, pch=15 )


# Create scatter plot for with GPA as y-axis and Exercise as X-value

#I count the occurence of each couple of values. Eg : number of time a=1 and b=1, number of time a=1 and b=2 etc...
AA <- xyTable(dataFrame$GPA,dataFrame$Activities)
 
#Now I can plot this ! I represent the dots as big as the couple occurs often
coeff_bigger <- 2
plot(AA$x , AA$y , cex=AA$number*coeff_bigger  , pch=16 , col=rgb(0,0,1,0.5) , xlab= "Gpa levels" , ylab="Exercise levels" , xlim=c(0,6) , ylim=c(0,6), main="Excercse levels per week VS GPA")
text(AA$x , AA$y , AA$number )
 
#Note : It's easy to make a function that will compute this kind of plot automaticaly :
represent_discrete_variable <- function(var1, var2 , coeff_bigger){
  AA=xyTable(var1,var2)
  plot(AA$x , AA$y , cex=AA$number*coeff_bigger  , pch=16 , col="chocolate1" , xlab= "value of a" , ylab="value of b" )
  text (AA$x , AA$y , AA$number )
}


########## Comparison for subquestion of exercise levels and energy levels ##############################

# Base histogram
hist(dataFrame$Energy.Level,  col=rgb(0,1,0,0.5), xlab="Levels of Exercise and Energy", 
     ylab="Response Count", main="Exercise per Week vs Energy Levels of Last Two Weeks", ylim = c(0,25))

# Second with add=T to plot on top
hist(dataFrame$Activities, col=rgb(1,0,0,0.5), add=T)

# Add legend
legend("topright", legend=c("Energy Levels","Exercise"), col=c(rgb(0,1,0,0.5), 
     rgb(1,0,0,0.5)), pt.cex=2, pch=15 )

```


### Exercise and GPA comparison

When looking at the scatter plot with variable size graphs above, we can see a uptick in GPA with more frequent exercise sessions throughout the week. 

### Exercise and Energl Levels
The data shows that students with a lower frequency of exercise sessions per week have increase energy levels.

###Hours of Credits and GPA

```{r}

ggplot(dataFrame, aes(x=Credits, y=GPA)) + 
    geom_point(
        color="orange",
        fill="#69b3a2",
        shape=21,
        alpha=0.5,
        size=6,
        stroke = 2
        ) + ggtitle("Comparision between Number of Credits and GPA")
```

###Between the STEM and Non-STEM Majors, which one of them has the higher GPA?

```{r}
STEMMaj_Data<- subset(dataFrame, Major == "Yes")
NONSTEMMaj_Data<- subset(dataFrame, Major == "No")

STEMMaj_Data %>% ggplot( aes(x=GPA)) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8)
NONSTEMMaj_Data %>% ggplot( aes(x=GPA)) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8)

```

